---
title: "Reproduction Instructions"
output: 
  github_document:
    html_preview: false
---

Note: In order to reproduce this analysis, you need the R package separation from GitHub, which you can install using `devtools::install_github("carlislerainey/separation")`. See below for the details about other R pacakges.

## `Makefile`

To reproduce the analysis, first clear all created files. Run:

```
make cleanALL
```

Then reproduce the analysis. Run:

```
make
```

### `make`ing

The `Makfile` describes the structure of the code and allows the user to reproduce the entire analysis or portions of it.

- `make all` or just `make` reproduces the entire analysis, including the simulations, which take about five hours.
- `make dag` reproduces a DAG that shows the structure of the dependencies in the `Makefile`.
- `make sims` reproduces the `.rds` files of the simulations (created by both `simulations.R` and `sample-size-simulations.R`) and saves them in the `simulations` directory. You can monitor the progress of `simulations.R` and `sample-size-simulations.R` with `simulations.log` and `sample-size-simulations.log`, respectively. This takes about five hours.
- `make simplots` reproduces the figures summarizing the simulations (Figures 2-5 and 8 from the main paper as well as figures for the appendix) and saves them in the `manuscript/figs` directory.
- `make ge` reproduces our re-analysis of George and Epstein (1992) and saves the figures to the `manuscript/figs` directory.
- `make weisiger` reproduces our re-analysis of Weisiger (2014) and saves the figures to the `manuscript/figs` directory.
- `make manuscript` recompiles the LaTeX manuscript `small.tex` into `small.pdf` and `small-appendix.tex` into `small-appendix.pdf`. It automatically handles the bibliography.
- `make readme` knits this document from the `.Rmd` file.
- `make computedvalues` knits `computed-values.pdf`, which creates tables of the numeric quantities reported in the text.

### Cleaning

You can clear any produced files from by the code by running `make clean*` and using any of the phonies above. For example,`make cleandag` removes the figure `makefile-dag.png`. To clean the entire directory, run `make cleanALL` (I put `ALL` in caps to remind myself of the consequences).

### `illustrate-bias-annotated.pdf`


We used a combination of ggplot and Apple Keynote to create Figure 1 `manuscript/figs/illustrate-bias-annotated.pdf`, which illustrates the source of the small sample bias. The R script `R/illustrate-bias.R` create the underlying plot `manuscript/figs/illustrate-bias.pdf`, but we added the annotations manually in Keynote.

## Additional Notes

- The simulations take about five hours. We've set them up to run in parallel on four clusters. You might speed this up with a change here.
- The code automatically stores the packages used in the last run in the file `session-info.txt`.
- There is no log file, but all the figures are created and saved in the `manuscript/figs` directory. All quantities reported in the manuscript are computed and/or reported in the file `computed-values.pdf`.

## R

We ran the analysis using the system below.

```{r}
R.version
```


## R Packages

In order to reproduce the analysis, several R packages, which you can install with the following code: 

```{r}
# list of packages on CRAN used in this project (exclusing base packages)
pkg <- c("brglm", 
         "brglm2", 
         "devtools",
         "doParallel", 
         "doRNG",
         "foreach", 
         "gridExtra", 
         "gridExtra", 
         "logistf", 
         "quantreg", 
         "scoring", 
         "texreg", 
         "tidyverse", 
         "xtable")
```

To install these packages, you can run the code above along with the command below.

```
install.packages(pkg, repos = "http://cran.rstudio.com")
```

You also need the package separation from GitHub, which you can install with the command below.

```
devtools::install_github("carlislerainey/separation")
```

We recommend using the latest version of each package, but the versions we used are saved to the file `package-versions.csv`.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)

devtools::package_info(pkgs = c(pkg, "separation"), dependencies = TRUE) %>%
  select(package, version = ondiskversion, date, source) %>%
  write_csv("package-versions.csv")
```

## Reproducing without `Make`

- To create `makefile-dag.png`, run the R script `makefile-dag.R`.
- To do the simulations for figures 2-5 and store them as `simulations/simulations.rds`, run the R script `R/simulations.R`.
- To create the figures based on the simulations above, run the R script `R/plot-simulations.R`. These figures are stored as `.pdf`s in `manuscript/figs`.
- To perform the sample size simulations for figure 8 and store them as `simulations/sample-size-simulations.rds`, run the R script `R/sample-size-simulations.R`.
- To create the figures based on the sample size simulations above, run `R/plot-sample-size-simulations.R`. These figures are stored as `.pdf`s in `manuscript/figs`.
- To reproduce the George and Epstein re-analysis, run the R script `ge-replication/R/analysis.R`. Figures are stored as `.pdf`s in `manuscript/figs`.
- To reproduce the Weisiger re-analysis, run the R script `weisiger-replication/R/analysis.R`. Figures are stored as `.pdf`s in `manuscript/figs`.
- To compile the manuscript and appendix, compile `manuscript/small.tex` and `manuscript/small-appendix.tex`, respectively, with pdftex and bibtex.
- To render the `README.md`, compile `README.Rmd` with knitr.
- To render the `computed-values.pdf`, compile `computed-values.pdf` with knitr.
